Spin canonical divisors of log stacky curves

### 2015-06-20 stacky_write_up/spin_cring.tex
wrote second inductive lemma

\begin{lem}
\label{sat_three_induction}
Let $\sx \rightarrow \sx'$ be a birational map of tame, separably rooted  log stacky curves $R \subset R'$ be the corresponding inclusion of canonical rings. Let $R_{D'} = R' = k[x]/I'$ satisfy $sat(Eff(D')) = 3.$ Let $D = \subhalf {e_i}$. There are generators $y_3,y_4,y_5$ of $R_{D+D'}= R$ over $R'$ so that $y_i$ lies in degree $i$ and has a pole of order $1$. \todo{Include the order}
Let $x_3,x_4,x_5$ be generators of $k[x]$ lying in degrees 3,4,5 respectively. \todo{Make sure to state some requirement on the ordering of the $x_i$, analogously to 8.4.1, but I don't see why this is important.} Then, letting $x_j$ range over all $x$ variables
\begin{align*}
	in_\prec(I) = in_\prec(I')k[x,y_3,\ldots, y_e,z_4,z_5] + \\
	\langle y_i y_j \mid 1 \leq i < j-3 \leq e \rangle +\\
	\langle y_i x_j \mid i > 3 \rangle +\\
	\langle z_i y_j \mid j > 3 \rangle  +\\
	\langle z_i x_j \mid 4 \leq i \leq 5\rangle + \\
	\langle z_iz_k \mid 4 \leq i \leq j \leq 5 \rangle 
\end{align*}

\end{lem}
\begin{proof}
We can see the following is a $k$ basis for $R$ over $R'$:
\begin{align*}
	\langle y_i^ay_{i+2}^b \mid a \geq 0,b\geq 0, 3 \leq i \leq e-2\rangle +\\
	\langle y_3^ax_3^b x_4^\epsilon x_5^{\epsilon'} \mid a \geq 0, b \geq 0(\epsilon,\epsilon') \in \{(0,0),(0,1),(1,0)\} \rangle +\\
	\langle y_3^az_4,y_3^bz_5 \mid a \geq 0, b \geq 0 \rangle\\
\end{align*}
Note that this is precisely the monomials which are not divisible by those in the claimed initial ideal. By the trick of projecting to the pole order by degree 2-dim lattice, we see this is a $k$ basis. We also see that $z_i,y_j$ are generators. \todo{Note that $z_5$ is chosen specially, and is not generic}

It only remains to show the initial ideal is as claimed. Note that once we show this, we will automatically obtain minimality, using the regrading to degree 1 trick.

To show this is the initial ideal, it suffices to exhibit such relations with the claimed leading terms in the initial ideal. Indeed, we will now give only leading order terms. We have relations starting with
\begin{align*}
	y_iy_j - y_{i-1}y_{j+1} + \cdots &\mid i < j-1 \\
	y_ix_4 - y_{i-2}y_3x_3 + \cdots & \mid i > 3\\
	y_ix_5 - y_{i-2}y_3x_4 + \cdots & \mid i > 3\\
	y_ix_3-y_3^3y_{i-6} + \cdots &\mid i > 7\\
	y_7x_3-y_3^2 z_4 + \cdots \\
	y_5x_3 - y_3z_5 + \cdots \\
	z_4^2 - y_3 z_5 + \cdots \\
	z_4z_5 - y_3^2x_3 + \cdots \\
 	z_5^2 - y_3^2 x_4 + \cdots \\
 	z_4 y_i - y_3^2 y_{i-2} + \cdots &\mid i > 3 \\
 	z_5 y_i - y_3y_{i-2}z_4 + \cdots &\mid i > 3 \\
 	z_i x_j - y_3 z_{i+j-3} + \cdots &\mid i+j \leq 8 \\
 	z_i x_j - y_3x_3x_{i+j-6} + \cdots &\mid i+j > 8
\end{align*}
\end{proof}
\todo{Something is a little fishy with this order, it seems something like: first grade by degree, then try to decrease the highest y, if you cant, try to decrease the highest z, possibly replace $z_5$ by $z_5+Cy_5$ to generic element?}

>>>>

\begin{prop}
\label{prop:sat_three_induction}
Let $\sx \rightarrow \sx'$ be a birational map of tame, separably rooted  log stacky curves $R \subset R'$ be the corresponding inclusion of canonical rings. Let $R_{D'} = R' = k[x]/I'= k[x_3,x_4,x_5,\ldots, x_m]/I'$ satisfy $sat(Eff(D')) = 3.$ Let $D = \frac{1}{3}P$. 
\begin{enumerate}
	\item[(a)] General elements  $y_i \in H^0(\sx,iL)$ for $i \in \{3,4,5\}$ satisfy $-\ord_P(y_i) = 1$ and any such choice of elements $y_3,y_4,$ and$ y_5$ minimally generate $R$ over $R'$.
	\item[(b)] Equip $k[y_3,y_4,y_5]$ with $grlex$ so that $y_3 \prec y_4 \prec y_5$ and equip the ring $k[y_3,y_4,y_5,x]$ with the block order so that $R = k[y_3,y_4,y_5,x]/I$. Let,
	\begin{align*}
	J &= gin_\prec(I')k[x,y_3,y_4,y_5] \\
	&+\langle y_i x_j \mid 4 \leq i \leq 5, 3 \leq j \leq m\rangle \\
	&+\langle y_iy_k \mid 4 \leq i \leq j \leq 5.\rangle 
\end{align*}
Then, $gin_\prec(I) = J$.
	\item[(c)] Any set of minimal generators for $I'$ together with any set of relations with leading terms as in (b) minimally generate $I$.
\end{enumerate}
\end{prop}
\begin{proof}
First, the elements $y_3,y_4,$ and $y_5$ are general and have $-\ord_P(y_i) = 1$ by Riemann Roch. \todo{Understand why/whether we need to cite lemma 5.4.7.} To complete the proof of (a) we only need show these elements generate $R$ over $R'$. We can first see that the following is a $k$ basis for $R$ over $R'$:
\begin{align}
\label{eqn:add_one_generator}
\begin{split}
	&\langle y_3^ax_3^b x_4^\epsilon x_5^{\epsilon'} \mid a \geq 0, b \geq 0,(\epsilon,\epsilon') \in \{(0,0),(0,1),(1,0)\} \rangle \\
	&+\langle y_3^ay_4,y_3^by_5 \mid a \geq 0, b \geq 0 \rangle
\end{split}
\end{align}
To see these generate $R'$ over $R$, note that $\dim R'_d - \dim_k R_d = \lfloor \frac{d}{3} \rfloor,$ by Riemann--Roch. So, it suffices to show that we have precisely $\lfloor \frac{d}{3} \rfloor,$ elements of degree $d$ in the claimed basis of ~\ref{eqn:add_one_generator}. Indeed, letting $a = \lfloor \frac{d}{3} \rfloor $ and $b = d \bmod 3$, we have that the elements
\begin{align*}
	x_3^{a-1}x_{3+b},y_3x_3^{a-2}x_{3+b},\ldots, y_3^ax_{3+b},y_3^ay_{3+b}
\end{align*}
are precisely $a$ elements, which are all independent as they have distinct pole orders at $P$. This completes part (a).

To show part (b), note that the generators in ~\ref{eqn:add_one_generator} are precisely a set of monomials which generate $k[x,y]/J$ over $k$. So, to show $J = in_\prec(I)$, it suffices to show that all generators of $J$ lie in $in_\prec(I)$. 
This follows, since there exist constants $A_{i,j} \in k$ for $4 \leq i \leq 5,3 \leq j \leq m,$ elements $B_1,B_2,B_3 \in k$ and elements $w_{i,j} \in R'$ so that the following linear combinations of elements lie in $R'$.

\begin{align*}
y_ix_j &- A_{i,j} y_{i-1}w_j & \text{ so that } 4 \leq i \leq 5,3 \leq j \leq m \text{ and } \deg w_j = \deg x_j + 1\\
y_4^2 &+ B_1 y_3 y_5 \\
y_4y_5 &+ B_2 y_3^2 x_3 \\
y_5^2 &+ B_3 y_3^2 x_4
\end{align*}
Of course, the initial terms of these elements are precisely the generators of $J$, completing (b).

Finally, (c) follows immediately from Lemma ~\ref{lem:minimal_quadratic}
\end{proof}
\todo{Something is a little fishy with this order, it seems something like: first grade by degree, then try to decrease the highest y, if you cant, try to decrease the highest z, possibly replace $z_5$ by $z_5+Cy_5$ to generic element?}
